---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data

* Unzip the file and store information on "data" variable 

```{r echo = TRUE}
unzip ("activity.zip", exdir = ".")
data <- read.csv("activity.csv", na.strings = "NA")
unlink("activity.csv")
```

* Convert dates into the correct format

```{r echo = TRUE}
data$date <- as.Date(as.character(data$date),"%Y-%m-%d")
```


## What is mean total number of steps taken per day?

For this calculations we will ignore the missing values in the dataset.


###1.Histogram of the total number of steps taken each day

```{r fig.width=4,fig.align='center',fig.height=4,echo = TRUE}
totalSteps <- aggregate(steps ~ date + steps, data, FUN=sum)
hist(totalSteps$steps, 
     main="Histogram for Total Number of Steps",
     xlab="Steps taken each day", border="blue", col="green",  breaks=15)

```


###2.The Mean and Median Total Number of Steps Taken Per Day

```{r echo = TRUE}
stepsMean <- round(mean(totalSteps$steps, na.rm = TRUE),0)
stepsMedian <- round(median(totalSteps$steps, na.rm = TRUE),0)
```
The Mean is `r stepsMean` and the Median is `r stepsMedian `.

## What is the average daily activity pattern?

###1.Time series plot of the 5-minute interval and the average number of steps taken, averaged across all days


```{r fig.width=4,fig.align='center',fig.height=4,echo = TRUE}
averageSteps <- aggregate(steps ~ interval + steps, data, FUN=mean)
plot(averageSteps, type = "l", 
     main="Daily Activity Pattern",
     xlab="Interval", col="green",  ylab = "Average Number of Steps")

```

###2.Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r echo = TRUE}

maxInterval <- averageSteps[which(averageSteps$steps == max(averageSteps$steps, 
                                                            na.rm = TRUE)), ]
```

The 5-miute interval with maximun number of steps in average is the interval `r maxInterval$interval`.


## Imputing missing values

Note that there are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data.

###1.Calculate and report the total number of missing values in the dataset

```{r echo = TRUE}

naTotal <- sum(is.na(data))
               
```

Amount of Missing Values In The Dataset is `r naTotal`.


###2. Strategy for filling in all of the missing values in the dataset. The strategy is the following: Use the mean of the 5-minute interval for missing value of that interval


a. Get Missing Values in the dataset
```{r echo = TRUE}
missingValues <- data[is.na(data$steps),]
```

b. For each missing interval value replace it with the correspondant mean

```{r echo = TRUE, eval=FALSE}

for(i in 1:nrow(missingValues)) {
    
    row <- missingValues[i,]
    
    #Get Mean Interval for missing interval
    aux <- averageSteps[which(averageSteps$interval == row$interval), 2]
    
    #Replace the mising step with the mean
    data$steps[which(is.na(data$steps) & data$interval == row$interval)] <- aux
}

```

###3.New dataset equal to the original but with the missing data filled in.

```{r echo = TRUE}

#Copy DataSet
data2 <- data

for(i in 1:nrow(missingValues)) {
    
    row <- missingValues[i,]
    
    #Get Mean Interval for missing interval
    aux <- averageSteps[which(averageSteps$interval == row$interval), 2]
    
    #Replace the mising step with the mean
    data2$steps[which(is.na(data2$steps) & data2$interval == row$interval)] <- aux

}

```

###4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. 


```{r fig.width=4,fig.align='center',fig.height=4,echo = TRUE}
totalSteps2 <- aggregate(steps ~ date + steps, data2, FUN=sum)
hist(totalSteps2$steps, 
     main="Histogram for Total Number of Steps",
     xlab="Steps taken each day", border="blue", col="green",  breaks=15)

```

```{r echo = TRUE}
stepsMean2 <- round(mean(totalSteps2$steps, na.rm = TRUE),0)
stepsMedian2 <- round(median(totalSteps2$steps, na.rm = TRUE),0)
```
The Mean is `r stepsMean` and the Median is `r stepsMedian `.


###Do these values differ from the estimates from the first part of the assignment? 

```{r echo = TRUE}
par(mfrow = c(1,2))

hist(totalSteps$steps, 
     main="Total Number of Steps With NA",
     xlab="Steps taken each day", border="blue", col="green", 
     ylim = c(0,30),  breaks=15)

hist(totalSteps2$steps, 
     main="Total Number of Steps With Mean Substitution",
     xlab="Steps taken each day", border="blue", col="green",  
     ylim = c(0,30), breaks=15)

```

###What is the impact of imputing missing data on the estimates of the total daily number of steps?


```{r echo = TRUE}

summary(data)

summary(data2)

```

3rd Quantile grew for NA substituted values


## Are there differences in activity patterns between weekdays and weekends?
